---
title: "DCF Valuation Model – Ferrari N.V."
author: "Pratiksha Madnurkar"
date: "2026-05-18"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Executive Summary

This project builds a Discounted Cash Flow (DCF) valuation model for Ferrari N.V. using R.

The objective is to estimate the intrinsic equity value of the company by forecasting future free cash flows and discounting them using the weighted average cost of capital (WACC).

The analysis includes:
- Revenue forecasting
- Free cash flow projections
- Terminal value estimation
- Enterprise value calculation
- Sensitivity analysis

# Business Objective

The purpose of this project is to estimate Ferrari’s theoretical fair value and evaluate how valuation changes under different assumptions.

Key business questions:
- What is Ferrari’s estimated intrinsic value?
- How sensitive is valuation to WACC and terminal growth assumptions?
- How do future cash flows contribute to enterprise value?

# DCF Assumptions

The model uses several core financial assumptions.

```{r}
library(tidyverse)

# === INPUTS — change these for any company ===
company     <- "Ferrari N.V. (RACE)"
revenue_base <- 5970    # last year revenue in €M
fcf_margin  <- 0.18    # free cash flow margin (18%)
growth_rate <- 0.12    # expected annual growth
wacc        <- 0.09    # weighted average cost of capital
terminal_g  <- 0.025   # terminal growth rate
years       <- 5       # projection horizon
net_debt    <- 550     # net debt in €M
shares_out  <- 182.8   # shares outstanding in millions
```

# Revenue & Free Cash Flow Forecast

```{r}
proj <- tibble(
  year    = 1:years,
  revenue = revenue_base * (1 + growth_rate)^(1:years),
  fcf     = revenue * fcf_margin,
  pv_fcf  = fcf / (1 + wacc)^(1:years)
)

proj
```

# Terminal Value Estimation

```{r}
terminal_fcf  <- tail(proj$fcf, 1) * (1 + terminal_g)
terminal_value <- terminal_fcf / (wacc - terminal_g)
pv_terminal   <- terminal_value / (1 + wacc)^years
```

# Enterprise Value Calculation

```{r}
enterprise_value <- sum(proj$pv_fcf) + pv_terminal
equity_value     <- enterprise_value - net_debt
price_per_share  <- equity_value / shares_out

cat("Enterprise Value: €", round(enterprise_value,0), "M\n")
cat("Implied Share Price: €", round(price_per_share,2), "\n")
```

# Sensitivity Analysis

The sensitivity table evaluates valuation changes under different WACC and terminal growth assumptions.

```{r}
wacc_range <- seq(0.07, 0.11, by=0.01)
tg_range   <- seq(0.01, 0.04, by=0.01)

sens_matrix <- outer(wacc_range, tg_range, Vectorize(function(w, g) {
  tv  <- (tail(proj$fcf,1)*(1+g)) / (w-g)
  ev  <- sum(proj$pv_fcf) + tv/(1+w)^years
  round((ev - net_debt)/shares_out, 0)
}))

rownames(sens_matrix) <- paste0(wacc_range*100, "% WACC")
colnames(sens_matrix) <- paste0("TG=", tg_range*100, "%")

print(sens_matrix)
```

# Key Insights

- Ferrari’s valuation is highly sensitive to discount rate assumptions.
- Terminal value contributes significantly to enterprise value.
- Higher growth assumptions materially increase implied share price.

# Conclusion

This project demonstrates practical corporate finance and valuation modeling skills.

Key competencies demonstrated:
- Financial forecasting
- DCF valuation
- Sensitivity analysis
- Corporate finance modeling
- Business valuation interpretation