---
title: "Econometrics Forecasting – GDP Analysis"
author: "Pratiksha Madnurkar"
date: "2026-05-18"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Executive Summary

This project analyzes quarterly GDP growth for Italy and Germany using econometric and time-series forecasting techniques.

The analysis includes:
- Eurostat data extraction
- GDP trend analysis
- ARIMA forecasting
- Regression analysis

# Business Objective

The purpose of this project is to evaluate macroeconomic trends and forecast future GDP growth.

Key business questions:
- How have Italy and Germany performed economically over time?
- Can future GDP growth be forecasted accurately?
- Does German GDP growth influence Italian GDP growth?


```{r}
library(eurostat)
library(tidyverse)
library(forecast)

search_eurostat("GDP") |> select(title, code) |> head(10)
gdp_raw <- get_eurostat("namq_10_gdp",
                        filters = list(
                          geo    = c("IT", "DE"),
                          na_item = "B1GQ",   # GDP
                          unit   = "CLV_PCH_PRE"  # % change previous quarter
                        )
)

```

# Data Cleaning

```{r}
gdp <- gdp_raw |>
  filter(time >= "2010-01-01") |>
  select(time, geo, values) |>
  drop_na()

glimpse(gdp)
```

# GDP Trend Analysis

```{r}
ggplot(gdp, aes(x=time, y=values, color=geo)) +
  geom_line(size=1) +
  geom_hline(yintercept=0, linetype="dashed", color="grey50") +
  scale_color_manual(values=c("IT"="#534AB7", "DE"="#185FA5"),
                     labels=c("IT"="Italy", "DE"="Germany")) +
  labs(title="Italy vs Germany: Quarterly GDP Growth (%)",
       subtitle="Source: Eurostat, 2010–2024",
       x=NULL, y="% change vs previous quarter",
       color="Country") +
  theme_minimal() +
  theme(legend.position="top")
```

# ARIMA Forecasting

```{r}
italy_gdp <- gdp |> filter(geo == "IT") |> pull(values)
ts_italy  <- ts(italy_gdp, frequency=4, start=c(2010,1))

arima_model <- auto.arima(ts_italy, seasonal=TRUE)
summary(arima_model)
```

# GDP Forecast

```{r}
fcast <- forecast(arima_model, h=8, level=c(80, 95))

autoplot(fcast) +
  labs(title="Italy GDP Growth Forecast — 8 Quarters",
       subtitle="ARIMA model with 80% and 95% confidence intervals",
       x=NULL, y="% growth") +
  theme_minimal()
```

# Regression Analysis

```{r}
gdp_wide <- gdp |>
  pivot_wider(names_from=geo, values_from=values) |>
  drop_na()

model <- lm(IT ~ DE, data=gdp_wide)
summary(model)

ggplot(gdp_wide, aes(x=DE, y=IT)) +
  geom_point(color="#534AB7", alpha=0.6) +
  geom_smooth(method="lm", color="#1D9E75", se=TRUE) +
  labs(title="Does Germany's GDP growth predict Italy's?",
       x="Germany GDP growth (%)", y="Italy GDP growth (%)") +
  theme_minimal()
```


# Key Insights

- GDP growth exhibited strong volatility during economic shocks.
- ARIMA models captured cyclical macroeconomic behavior effectively.
- German GDP growth demonstrated moderate predictive influence on Italian GDP growth.

# Conclusion

This project demonstrates practical econometric and forecasting skills.

Key competencies demonstrated:
- Time-series forecasting
- ARIMA modeling
- Regression analysis
- Macroeconomic analysis
- Data visualization