---
title: "FP&A Dashboard"
author: "Pratiksha Madnurkar"
date: "2026-05-18"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Executive Summary

This project analyzes Ferrari S.p.A.'s financial performance from 2015–2025 using R.
The objective is to simulate a real-world FP&A workflow including:
- Financial KPI tracking
- Margin analysis
- Revenue trend visualization
- Year-over-year growth analysis
- Interactive dashboarding

The dashboard was developed using R, ggplot2, Plotly, and Shiny.

# Business Objective

The purpose of this project is to evaluate Ferrari’s financial growth and profitability trends over time.

Key business questions:
- How has revenue evolved over the last decade?
- Are profitability margins improving?
- What are the company’s major growth trends?
- How can financial KPIs be visualized interactively for management reporting?

```{r}
library(tidyverse)
library(scales)
library(readxl)

financials <- read_xlsx("C:/Users/prati/OneDrive/Documents/Jobs/Projects/Financials.xlsx")
glimpse(financials)
```


# Financial KPI Engineering

Additional KPIs were created to support FP&A analysis.

These include:
- EBITDA Margin
- Net Profit Margin
- Year-over-Year Revenue Growth

```{r}
financials <- financials |>
  mutate(
    ebitda_margin = round(EBITDA/`Net Revenue` * 100, 1),
    net_margin    = round(`Net Income`/`Net Revenue` * 100, 1),
    yoy_growth    = round((`Net Revenue`/lag(`Net Revenue`) - 1) * 100, 1))
```


# Revenue Trend Analysis

The chart below visualizes Ferrari’s revenue growth over time.

```{r}
ggplot(financials, aes(x = Year, y = `Net Revenue`)) +
  geom_col(fill = "#534AB7", width = 0.6) +
  geom_text(aes(label = paste0(round(`Net Revenue`/1000,1), "B")),
            vjust = -0.5, size = 3.5, color = "#26215C") +
  labs(title = "Ferrari Revenue 2015–2025 (€M)",
       x = "Year", y = "Revenue (€M)") +
  theme_minimal() +
  scale_y_continuous(labels = comma)
```


## Key Insight:
Ferrari demonstrates a strong upward revenue trajectory over the observed period, indicating sustained demand growth and pricing power.

# Profitability Analysis

The following visualization evaluates profitability margins over time.

```{r}
ggplot(financials, aes(x = Year, y = ebitda_margin)) +
  geom_line(size = 1.2, color = "#1D9E75") +
  geom_point(size = 2.5, color = "#534AB7") +
  labs(
    title = "EBITDA Margin Trend",
    x = "Year",
    y = "EBITDA Margin (%)"
  ) +
  theme_minimal()
```

## Key Insight:
The EBITDA margin trend suggests operational efficiency improvements and strong cost management.

# Year-over-Year Growth
```{r}
ggplot(financials, aes(x = Year, y = yoy_growth)) +
  geom_col(fill = "#185FA5") +
  labs(
    title = "Year-over-Year Revenue Growth",
    x = "Year",
    y = "Growth (%)"
  ) +
  theme_minimal()
```


## Key Insight:
Revenue growth accelerated significantly in post-pandemic years, reflecting recovery in luxury automotive demand.

# Interactive Dashboard

The project also includes an interactive dashboard built using Shiny and Plotly.

Key dashboard features:
- Dynamic metric selection
- Interactive visualizations
- Financial table outputs
- User-friendly management reporting interface

```{r}
library(shiny)
library(tidyverse)
library(plotly)
library(DT)
```
```{r}
ui <- fluidPage(
  titlePanel("Ferrari S.p.A. — Financial Performance Dashboard"),
  sidebarLayout(
    sidebarPanel(
      selectInput("metric", "Select metric:",
                  choices = c("Revenue"="revenue",
                              "EBITDA"="ebitda",
                              "Net Income"="net_income"))
    ),
    mainPanel(
      plotlyOutput("chart"),
      br(),
      DTOutput("table")
    )
  )
)
```

```{r}
server <- function(input, output) {
  output$chart <- renderPlotly({
    p <- ggplot(financials, aes(x=year, y=.data[[input$metric]])) +
      geom_col(fill="#534AB7") +
      geom_line(color="#1D9E75", size=1) +
      theme_minimal() +
      labs(x="Year", y="€M")
    ggplotly(p)
  })
  output$table <- renderDT({
    datatable(financials, options=list(pageLength=5))
  })
}
```
```{r}
shinyApp(ui, server)
```

#Forecating

```{r}
library(forecast)

# Simple linear forecast for next 2 years
rev_ts <- ts(financials$`Net Revenue`, start=2015)
model  <- auto.arima(rev_ts)
fcast  <- forecast(model, h=3)

autoplot(fcast) +
  labs(title="Revenue forecast 2026-2027",
       x="Year", y="Revenue (€M)") +
  theme_minimal()
```

# Conclusion

This project demonstrates practical FP&A and financial analytics skills using R.

Key competencies demonstrated:
- Financial KPI analysis
- Data visualization
- Interactive dashboard development
- Business performance interpretation
- Financial storytelling

The project simulates a real-world corporate finance reporting workflow commonly used in FP&A teams.