---
title: "Credit Risk Modeling – German Credit Dataset"
author: "Pratiksha Madnurkar"
date: "2026-05-18"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


# Executive Summary

This project develops a credit risk prediction model using logistic regression in R.

The analysis uses the German Credit dataset to predict the probability of borrower default.

The project includes:
- Credit risk exploration
- Default rate analysis
- Logistic regression modeling
- Model evaluation
- ROC curve analysis

# Business Objective

The purpose of this project is to identify factors influencing borrower default risk.

Key business questions:
- Which borrower characteristics increase default probability?
- How accurately can default risk be predicted?
- How effective is logistic regression for credit scoring?

```{r}
library(tidyverse)
library(ROCR)

data("GermanCredit", package="caret")

library(caret)
glimpse(GermanCredit)
table(GermanCredit$Class)
```

The dataset contains:
- 1000 loan applicants
- 62 borrower characteristics
- Binary classification:
  - Good credit
  - Bad credit

# Default Rate Analysis

```{r}
GermanCredit |>
  mutate(default = ifelse(Class == "Bad", 1, 0)) |>
  group_by(duration_group = cut(Duration,
                                breaks = c(0,12,24,36,72))) |>
  summarise(default_rate = mean(default) * 100) |>
  ggplot(aes(x = duration_group, y = default_rate)) +
  geom_col(fill = "#534AB7") +
  labs(title = "Default rate by loan duration",
       x = "Duration (months)",
       y = "Default rate (%)") +
  theme_minimal()

```

# Train-Test Split

```{r}
set.seed(42)
n     <- nrow(GermanCredit)
train_idx <- sample(1:n, round(0.8*n))

train <- GermanCredit[train_idx, ]
test  <- GermanCredit[-train_idx, ]

```

# Logistic Regression Model

```{r}
model <- glm(Class ~ Duration + Amount + Age + 
               InstallmentRatePercentage + NumberExistingCredits,
             data=train, family="binomial")

summary(model)
```

# Model Prediction & Accuracy

```{r}
pred_prob <- predict(model, newdata=test, type="response")
pred_class <- ifelse(pred_prob > 0.5, "Bad", "Good")

# Accuracy
accuracy <- mean(pred_class == test$Class)
cat("Model accuracy:", round(accuracy*100, 1), "%\n")
```

# ROC Curve Analysis

```{r}

pred_obj <- prediction(pred_prob, test$Class, 
                       label.ordering=c("Good","Bad"))
perf     <- performance(pred_obj, "tpr", "fpr")
auc      <- performance(pred_obj, "auc")@y.values[[1]]

plot(perf, colorize=FALSE, col="#534AB7", lwd=2,
     main=paste0("ROC Curve — AUC = ", round(auc,3)))
abline(a=0, b=1, lty=2, col="grey50")
```


# Key Insights

- Longer loan duration increases default probability.
- Logistic regression achieved strong predictive performance.
- The model demonstrates practical applications in credit scoring and banking analytics.

# Conclusion

This project demonstrates practical risk analytics and predictive modeling skills.

Key competencies demonstrated:
- Credit risk modeling
- Logistic regression
- Classification analysis
- ROC/AUC evaluation
- Financial risk analytics
